---
title: "SoccerFouls"
Rachel Lawrence
---

Do fouls matter in women's college soccer? In this assignment, you'll be working with a dataset that contains information about soccer matches from last season. You'll load the data, perform some basic statistical analysis, and create a visualization to explore the relationship between fouls and the goal differential (and thus the outcome). Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load the soccer dataset from https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/refs/heads/main/data/ncaa_womens_soccer_matchstats_2024.csv

```{r}
matches <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/refs/heads/main/data/ncaa_womens_soccer_matchstats_2024.csv")
```


### Task 3: Create a new column called `goal_diff` that calculates the difference between the two teams' goals.

```{r}
matches <- matches |> mutate(goal_diff = team_score - opponent_score)
```


### Task 4: Use `summarize()` and the `cor()` function to calculate the correlation between fouls and goal differential, using the pearson method.

```{r}
matches |> summarize(correlation = cor(fouls, goal_diff, method="pearson"))
```
*Correlation is basically nothing*


### Task 5: Create a scatterplot of fouls and goal differential, using `geom_point()`

```{r}

matches |> 
  ggplot() +
  geom_point(aes(x=goal_diff, y=fouls)) +
  geom_smooth(aes(x=goal_diff, y=fouls), method="lm")

```
*Flat line? Effecitvely no relation whatsoever*

### Task 6: Describe the results of both the correlation calculation and the scatterplot.

There is virutally no correlation between fouls and goal differential. 

### Task 7 (BONUS! YAY!): Choose another variable and repeat. 


```{r}
matches |> summarize(correlation = cor(team_score, goal_diff, method="pearson"))
matches |> 
  ggplot() +
  geom_point(aes(x=goal_diff, y=team_score)) +
  geom_smooth(aes(x=goal_diff, y=team_score), method="lm")
```

