---
title: "JOUR405: Statistics for Journalists"
subtitle: "Midterm Exam"
name: "Rachel Lawrence"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

## Load the tidyverse

```{r}
library(tidyverse)
```

---

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible, the compliance score, and the grade.

### Task 1: Read the data (2 points)

```{r}
read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
health_inspection <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```

### Task 2: Calculate the mean, standard deviation, minimum, and maximum of compliance scores. Save the results in a new dataframe called `compliance_stats`. (2 points for each)

```{r}
compliance_stats <- health_inspection |> 
  summarize(mean_compliance = mean(compliance_score),
            sd_compliance = sd(compliance_score),
            min_compliance = min(compliance_score),
            max_compliance = max(compliance_score))
```

### Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a descriptive title and label the x and y axes using `labs()`. (10 points)

```{r}

ggplot(compliance_stats, aes(x = compliance_score)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  geom_vline(xintercept = mean_compliance, color = "red", linetype = "dashed", linewidth = 1 ) +
  facet_wrap(~sample_size) +
  labs(
    title = "Compliance Scores in Health Inspections Across Montgomery County",
    subtitle = "Red line = mean compliance score.",
    x = "Compliance Score",
    y = "Count",
    caption = "Source: 2024 Cooperative Election Study"
  ) +
  theme_minimal()

```

### Task 4: Write 2–3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What is newsworthy, and what would your next reporting step be? (5 points)

**Answer:**
The standard deviation shows that scores were mstly contained within + or - 15 of the mean, which also shows that variation between scores was wide. (My histogram isn't work). As a reporter, I would want to visit towns that fell within one, two, and three sds of the mean, and see how different they really look. For example, why did one area do very poorly while another close by did very well?
---

## Part 2: Daily Transit Ridership — Z-Scores and Sampling (30 points)

Your editor wants you to identify unusual days in WMATA's daily ridership data that might be worth investigating as story leads. The dataset contains daily bus and rail totals. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv`

### Task 5: Load the WMATA ridership data. (2 points)

```{r}
wmata_daily <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```

### Task 6: Using `summarize()`, calculate the mean and standard deviation for both bus and rail ridership. Save the results in a dataframe called `wmata_stats`. (5 points)

```{r}
wmata_stats <- wmata_daily |> 
  summarize(mean_rail = mean(rail),
            sd_rail = sd(rail),
            mean_bus = mean(bus),
            sd_bus = sd(bus))
```

### Task 7: Using `mutate()`, add two new columns to `wmata_daily`: `rail_zscore` (the z-score for rail ridership) and `bus_zscore` (the z-score for bus ridership). Save the result as `wmata_zscores`. (8 points)

```{r}
wmata_zscores <- wmata_daily |>
   mutate (mean_rail = mean(rail),
    sd_rail = sd(rail),
    rail_zscore = (rail - mean_rail) / sd_rail, 

    mean_bus = mean(bus),
    sd_bus = sd(bus),
    bus_zscore = (bus - mean_bus) / sd_bus)

```

### Task 8: Filter `wmata_zscores` to show only days where the absolute value of `rail_zscore` is greater than 2 (unusually high or low rail ridership). Arrange by the absolute value of the z-score, with the most extreme first. How many unusual days are there? (8 points)

```{r}
unusual_rails <- wmata_zscores |>
  filter(abs(rail_zscore) > 2) |>
  arrange(desc(abs(rail_zscore)))
```

### Task 9: Using `set.seed()`, take a random sample of 91 days from `wmata_daily` (roughly 25% of the data). Calculate the mean and standard deviation for bus and rail ridership in the sample. How do the sample statistics compare to the population statistics you calculated in Task 6? Write 1–2 sentences comparing the values and explaining what this suggests about random sampling. (4 points)

```{r}
set.seed(42)
sample_91 <- wmata_daily |> sample_n(91)
sample_91 <- sample_91 |>
    summarize(mean_91 = mean(sample_91), 
            sd_91 = sd(sample_91))
  
```

### Task 10: Write 2–3 sentences: What do the unusual days in your z-score analysis suggest as story leads? (3 points)

The unusual rail scores suggest a few things, specifically that holidays and the first days back from winter breaks are popular to ride the metro on. All 5 of these days apply to that group. (Christmas, Thanksgiving, New Years Day, and the first Monday back after New Years Day). I would write that the metro provides people easy travel on days where traffic on the road is probably the highest. 
---

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to find out which counties have the highest rates. The dataset contains car theft counts for Maryland counties in 2023 along with population figures. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv`

### Task 11: Load the car thefts data. (2 points)

```{r}
md_car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

### Task 12: Using `mutate()`, add a column called `theft_rate` that calculates the car theft rate per 10,000 residents for each county. Save the result back to `md_car_thefts`. In 1–2 sentences, explain why a per-capita rate is more useful than raw counts for this comparison. (8 points)

```{r}
md_car_theft_rate <- md_car_thefts |>
  mutate(theft_rate = (thefts_2023 / population) * 10000)
```
I believe a per captia rate is more useful than the raw counts because it makes the statistics relative to the size of the county. Obviously, counties with more people, like Baltimore County, are going to have more thefts. The per-capita rate shows the crime based on the number of people in the county, which is how much smaller counties, like St. Mary's County, are further down on the list. 


### Task 13: Calculate the median car theft rate and the total number of car thefts statewide. Show the counties that have rates above the median. (7 points)

```{r}
md_car_theft_rate |>
  summarize(median = median(theft_rate))
```
The counties that have a rate above the median are Wicomico, Cecil, Charles, Washington, Anne Arundel, Montgomery, Charles, Dorchester, Baltimore, PG County, St. Mary's and Baltimore City. 

### Task 14: Write 2–3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What is the lede of a story based on your findings? (8 points)

These calculation show that car theft a) happens in every county, and that b) half of the counties experience more than 10 car thefts per 10000 people. I think this also shows that car theft per capita doesn't have much to do with the population of the city, as even the most poulated city doesn't have the highest theft-per-capita rate. 

LEDE: In half of Maryland's counties, at least 10 cars per 10,000 people are stolen every year, a study shows. 
_____________________________________________________________________________________________________________________________________________________________
## Part 4: Data Analysis Scenario (20 points)

A city parks department releases data showing that its 12 community swimming pools average 187 visitors per day during the summer season, with a standard deviation of 42 visitors. Your editor suspects that a handful of pools are dramatically underperforming and wants a story. You also learn that the department's figures are based on a random sample of 60 days per pool rather than a complete count.

Write 3–4 sentences for each of the following questions:

### Question A: Given that the mean is 187 and standard deviation is 42, which statistical measures would you prioritize using and why? (7 points)

I would use an analysis of the z-scores here to determine if there are actually pools underperforming more than others or if all of the pools are simply on a attendance decline. The z-score would more comprehensively show if the values are concerns for investigation, but if they fall within 2, I don't think the editor would be correct. 

### Question B: How would you use z-scores to identify which pools are genuinely underperforming, and why is that more useful than simply ranking pools by raw attendance? (7 points)

I would look at the z-scores that were above and below 3 to determine what pools are the mose "unusually" attended or unattended. I think this is more accurate than just raw attendance because the z-score takes into consideration regular attendance. For example, 20 attendees could be a very popular day for one pool but an under-attended day for another. 

### Question C: The department collected its attendance figures using a random sample of 60 days per pool rather than tracking every day. What questions would you ask about the sample before relying on these figures for a story? What concerns would you raise in your story? Use statistical terminology where appropriate. (6 points)

I would ask about what days the sample ended up examining. There's a chance that 20 of those days were the coldest of the season, which would alter the data because people don't usually flock to pools when it's coldest out. I would also have concerns that the days included holidays or days with early/late opening, as people may not want to go on a day where the pool closes early anyway. I would raise the concern that the sample should be colleted with more consideration of those external factors, and not purely random. I'd recommend they use a stratified sample instead. 
---

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your choices

Good luck!
