---
title: "HW11: Education Level and Survey Weighting"
author: Rachel Lawrence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

# Introduction

When a poll or survey gets published, the results almost always include a note about sample size and methodology. One thing you'll often see: "results have been weighted to reflect the U.S. adult population." But what does weighting actually *do* to the numbers — and does it matter for your story?

In this assignment, you'll compare weighted and unweighted results from a 2020 survey on voting intentions, focusing on how education level relates to whether someone planned to vote.

## The Data

```{r load-data}
nonvoters_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405_files/refs/heads/main/nonvoters_data.csv")
glimpse(nonvoters_data)
```

Key variables:

- `weight`: Each respondent's survey weight. A weight of 1.0 means they count as exactly one person. A weight of 2.0 means their response counts double (they represent an undersampled group). A weight of 0.5 means their response counts half as much (they represent an oversampled group).
- `Q21`: Voting intention — 1 = Yes, 2 = No, 3 = Unsure
- `educ`: Education level — College, Some college, High school or less

## Task 1: Examine the Weights (4 points)

Before comparing weighted and unweighted results, let's understand what the weights look like.

```{r examine-weights}
summary(nonvoters_data$weight)

ggplot(nonvoters_data, aes(x = weight)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Survey Weights",
    x = "Weight Value",
    y = "Number of Respondents"
  )
```

**Question 1:** Most weights cluster near 1.0, but some are much higher or lower. What does a weight of 0.2 tell you about that respondent's group in the sample? What about a weight of 3.5? Why might a pollster assign such extreme weights?

I think a really low weight means that a certain group was extrememly over represented, perhaps due to the nature of the poll, and counting them all the same would result in results that don't make sense with the population. The opposite applies to a really high weight--that group in the poll was underrepresented, and it wouldn't align with the population. I think a pollster would apply extreme weights to make sure that the poll can accurately represent the population.

## Task 2: Education Distribution in the Sample (3 points)

```{r education-distribution}
nonvoters_data |>
  count(educ) |>
  mutate(percentage = round(n / sum(n) * 100, 1))
```

**Question 2:** College graduates are overrepresented in most surveys. Does that appear to be true here? Why might college graduates be more likely to respond to a survey than people with less education?

I think college grads are slightly overrepresented in this survey, as I would expect it to be about a third of respondents. I think college grads might be more likely to respond to a survey because it's more of a part of daily life. At school, we're always being asked to do a survey or help a psych class or create surveys, but that was less common (at least for me) outside of campus. 

## Task 3: Unweighted Voting Intentions by Education (4 points)

This is what the results look like if you treat every respondent equally — no weights applied.

```{r unweighted}
unweighted_by_education <- nonvoters_data |>
  filter(!is.na(Q21), Q21 > 0, !is.na(educ)) |>
  group_by(educ, Q21) |>
  summarize(count = n(), .groups = "drop_last") |>
  mutate(total = sum(count),
         percentage = count / total * 100) |>
  ungroup()

unweighted_summary <- unweighted_by_education |>
  pivot_wider(
    id_cols = educ,
    names_from = Q21,
    values_from = percentage,
    names_prefix = "pct_"
  ) |>
  rename("Yes (%)" = pct_1, "No (%)" = pct_2, "Unsure (%)" = pct_3)

unweighted_summary
```

**Question 3:** Based only on these unweighted numbers, write a one-sentence headline for a story about education and voting intentions.

College graduates are more likely to vote than those with little or not college experience, a survey shows. 

## Task 4: Weighted Voting Intentions by Education (2 points)

Now apply the weights. Instead of counting respondents, we sum their weights — so someone with a weight of 2.0 contributes twice as much as someone with a weight of 1.0.

```{r weighted}
weighted_by_education <- nonvoters_data |>
  filter(!is.na(Q21), Q21 > 0, !is.na(educ)) |>
  group_by(educ, Q21) |>
  summarize(weighted_count = sum(weight), .groups = "drop_last") |>
  mutate(weighted_total = sum(weighted_count),
         weighted_percentage = weighted_count / weighted_total * 100) |>
  ungroup()

weighted_summary <- weighted_by_education |>
  pivot_wider(
    id_cols = educ,
    names_from = Q21,
    values_from = weighted_percentage,
    names_prefix = "pct_"
  ) |>
  rename("Yes (%)" = pct_1, "No (%)" = pct_2, "Unsure (%)" = pct_3)

weighted_summary
```

## Task 5: Visualize the Difference (4 points)

```{r visualization}
viz_data <- bind_rows(
  unweighted_by_education |>
    filter(Q21 == 1) |>
    mutate(Type = "Unweighted") |>
    select(Type, educ, percentage),
  weighted_by_education |>
    filter(Q21 == 1) |>
    mutate(Type = "Weighted",
           percentage = weighted_count / weighted_total * 100) |>
    select(Type, educ, percentage)
)

ggplot(viz_data, aes(x = educ, y = percentage, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)),
            position = position_dodge(width = 0.9), vjust = -0.5, size = 3) +
  labs(
    title = "Weighted vs. Unweighted: Share Planning to Vote, by Education",
    x = "Education Level",
    y = "% Saying Yes"
  ) +
  scale_fill_manual(values = c("Unweighted" = "#2166ac", "Weighted" = "#d6604d")) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Question 4:** Which education group shows the biggest shift between weighted and unweighted results? In percentage points, how large is that shift? Does weighting change the story, or does it just change the exact numbers?

Those with a high school diploma or less had the biggest shift in % saying yes between weighted and unweighted results. The shift is about 3%. I think this doesn't really change the story, but the exact numbers are definitely affected. 

## Task 6: The UMD Belonging Survey (8 points)

Earlier in class this week, we looked at the 2024 UMD Belonging and Community Survey. That survey had a 23% response rate and the report acknowledged "statistically significant differences" between the sample and the UMD population — but applied no weighting to correct for them.

**Question 5:** Based on what you've learned in this assignment, write 2-3 paragraphs as if you were a reporter evaluating the UMD survey before writing a story based on it. Address: (1) what the lack of weighting means for how you'd use the results, (2) which groups in the UMD survey appeared most underrepresented based on Figures 1 and 2, and (3) what question you would ask the survey's authors before publishing.

The lack of weighting basically means that the sample isn't quite representative of the population it represents. The results, therefore, really only apply to people who answered the questions. I would use these results sparingly, focusing mostly on the categories less related to race, gender and sexualtiy, especially "Considered Departure and Reasons for Staying." The most important takeaways from this category, in my opnion, have less to do with the gender, ethnic and sexual identiy of the repsondants, but rather how each individual thought. 

Basically anyone who isn't White, Black or Asian/Asian American is underrepresented, especially Pacific Islander/Native Hawaiian and Native American/Alaska Native students. This is mostly because of the non-specificity of the ambiguous "other" category. Students have the option to pick multiracial, but does that abandon one idenity? Are students who are Asian and Jewish forced to pick multi-racial, even if they are only of Asian descent? There are too many "what ifs," meaning a lot of students are not represented in the survey. 

I would ask the pollsters the purpose of weighting gender identity and race differently. In the race category, categories that have 0% of the sample, and subsequenty were weighted at 0%. However, "gender exapnsive" students also had 0% of the sample, but got 3% of the sample size. To me, that doesn't make sense. 
