---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2025.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

What stuck out to me the most is that the highest paid person in this White House is simply and "advisor." From a reporting standpoint, I would wonder how precarious this person's positoin is, and how important their "advice" is to whoever they advise (also not listed). Additionally, one of the least paid people, making 0 dollars per annum is also listed as an Advisor. An issue about this data is that "advsor" could mean literally anything. From a reporting standpoint, it'd be iportant to know what they do or who they advise, because at face value, it looks like two people are doing the same job, and one is making 200k+, while the other makes nothing. 

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
comp <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2025.csv")

comp |>
  summarize(
    mean_SALARY = mean(SALARY),
)
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
comp |>
  summarize(
    mean_SALARY = mean(SALARY),
    median_SALARY = median(SALARY),
    min_SALARY = min(SALARY),
    max_SALARY = max(SALARY)
)
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The median is greater than the mean, meaning the data is skewed right. I believe this means that most people make more money than the average number. The minimum and maximum salaries generally show the range of salaries across the White House staff. I'd consider those who make 0 dollars to be outliers because the next lowest salaries is almost 60K dollars. I believe that the several 0-dollar salaries may have affected the mean. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

My story idea would be the difference between advisors' salaries. Im very crusious to know why some make no money and some make the most. I would start by asking who they advise, if advising is their main source of income, how long they work, etc. 
