---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |> mutate(violent_rate_2019 = violent2019/pop2019 *1000, 
                    property_rate_2019 = property2019/pop2019 *1000,
                    violent_rate_2020 = violent2020/pop2020 *1000,
                    property_rate_2020 = property2020/pop2020 *1000)

```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

The city with the lowest amount of violent and property crime rates Eagle Harbor, which also only has 65 people.
The city with the highest crime rates is Ocean City, Maryland. I think this is because Ocean City is a huge tourist city in Maryland, and not all crimes committed there are by Ocean City residents. 

A pattern I noticed was that in 2019 and 2020, there are 20 cities where no violent or property crimes occur. 15 of those cities have less than 1000 people. Originally, I thought this to be a sign of a correlation bewteen population and crime rates, but on the flip side, Baltimore, the most populated city on the list, doesn't have the highest crime rate, so that doesn't quite apply. 

Like I mentioned about Ocean City, Baltimore also has a high tourist and non-reisdent population. This leads me to hypothesize that most crimes committed in these two areas are not committed by city residents. 

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities |> mutate(pct_change = (`violent2020`- `violent2019`)/`violent2019` * 100,
                                        pct_change = (`property2020` - `property2019`)/`property2019` * 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

There are few patterns in this data as it's truly all over the place. With that, I think the outliers are the most interesting, as well as comparing similar cities in the same county. In some smaller counties, like Calvert county, one town as vast increase in crime while the other has a significant decrease. 

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

The only two municipalities in Somerset County, Maryland experienced extremely opposite crime trends between 2019 and 2020. 

Despite only an increase of 16 residents between 2019 and 2020, Crisfield saw a a 318.2% increase in violent and property crimes, the largest increase in the state of Maryland. Princess Anne, the only other town in Somerset County, lost 12 residents over the year. However, the city, saw a 17.9% decrease over the year instead. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
