RACHEL LAWRENCE
JOUR 405
Final Project

My hypothesis: Best Picture winners at the Oscars don't reflect the opinions of the millions of non-critic audience members.  

```{r}
library(tidyverse)
```

```{r}
imdb_rt <- read_csv("imdb_rt_oscars.csv")
```

April 22nd: The Tests I Can Run

To prove my hypothesis, there are a lot of tests that would be valuable. To start, I'd simply examine the mean and quartiles of the scores for IMBD over those of Rotten Tomatoes. 

```{r}
summary(imdb_rt$`IMDb_Rating (Fan Rating)`)
```

```{r}
summary(imdb_rt$`RT Rating (Critic)`)
```

On average, movie critics (via Rotten Tomatoes) gave movies 1.296 higher of a rating than fans. Obviously, critics were harsher with a lower minimum rating of only 4.2 out of 10, where fans were more generous to Best Picture winners they didn't enjoy. 


I used a scatterplot next to visualize a potential relationship between Fan Ratings and Critic Ratings. I included a line of best fit, but it was basically useless. There is virtually no relationship between the two. 

```{r}
imdb_rt |>
  ggplot(aes(x = `IMDb_Rating (Fan Rating)`, y = `RT Rating (Critic)`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "pink")
```




