RACHEL LAWRENCE
JOUR 405
Final Project

My hypothesis: Best Picture winners at the Oscars don't reflect the opinions of the millions of non-critic audience members.  

```{r}
library(tidyverse)
```

```{r}
imdb_rt <- read_csv("imdb_rt_oscars.csv")
```

April 22nd: The Tests I Can Run
________________________________
To prove my hypothesis, there are a lot of tests that would be valuable. To start, I'd simply examine the mean and quartiles of the scores for IMBD over those of Rotten Tomatoes. 

```{r}
summary(imdb_rt$`IMDb_Rating (Fan Rating)`)
```

```{r}
summary(imdb_rt$`RT Rating (Critic)`)
```

On average, movie critics (via Rotten Tomatoes) gave movies 1.296 higher of a rating than fans. Obviously, critics were harsher with a lower minimum rating of only 4.2 out of 10, where fans were more generous to Best Picture winners they didn't enjoy. 


Next, to ensure there's no relation between the two, I ran a correlation model. 

```{r}
imdb_rt |> summarize(correlation = cor(`IMDb_Rating (Fan Rating)`, `RT Rating (Critic)`, method="pearson"))
```

This definetly surprised me. I was expected more of a disconnect between the two but the correlation between fan and critic ratings is decently strong. 
